---
title: "Tidying"
author: "Monica"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(readxl)
library(janitor)
library(tidyverse)

```


```{r}
inverts <- read_excel(here("data", "inverts.xlsx"))
```

## `pivot_longer()`
 Use single colun (:) to indicate a series of sequential columns. A:D

```{r}

inverts_long <- pivot_longer(data = inverts, 
                             cols = '2016':'2018',
                             names_to = "year",
                             values_to = "sp_count")

```

What if we want to get it back into wide format?

```{r}

inverts_wide <- inverts_long %>% 
  pivot_wider(names_from = common_name,
              values_from = sp_count)

```

### `janitor::clean_names()`

```{r}
inverts_wide <- inverts_wide %>% 
  clean_names()

```

## `tidyr.unite()` to combine information from columns

Example: merge site code and year into a single column called "site_year" 
```{r}
inverts_unite <- inverts_long %>% 
  unite(col = "site_year",
        c(site, year),
        sep = "_")


```


Create a new object called 'inverts_moyr' starting from inverts_long that unites the month and year columns into a SINGLE column named 'mo_yr' using a slash "/" as the separator

```{r}

inverts_moyr <- inverts_long %>% 
  unite(col = "mo_yr",
        c(month, year),
        sep = "_")

```

### `tidyr::separate()`

```{r}

inverts_sep <- inverts_unite %>% 
  separate(col = site_year, 
           into = c("my_site", "my_year"))
```

### `str_replace()`

```{r}

ca_abbr <- inverts %>% 
  mutate(common_name = 
           str_replace(common_name,
                       pattern = "california",
                       replacement = "CA"))

```

Example: In the common_name column, you want to replace spiny with awesome

```{r}
awesome_lobsters <- inverts %>% 
  mutate(common_name = 
           str_replace(common_name,
                       pattern = "spiny",
                       replacement = "awesome"))
```




### `stringr` + `mutate()`
Example using built in trees set with mutate() to add 5 to all tree heights
```{r}

trees_new <- trees %>% 
  mutate(height_update = Height + 5)


```

